Comparative study of Dask and Spark for neuroimaging pipelines\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Natural Sciences and Engineering Research Council}

### 2019-06-20 latex/paper.tex
Modified results section for incrementation and fixed typos

% Approach
We propose to use Dask since it provides native Python parallelism to pipelines while
providing support to familiar API from Python's scientific ecosystem. Unfortunately,
there are few comparisons between Spark and Dask; especially in neuroimaging.
Moreover, these studies were done when Dask was still an immature framework which
makes those studies unfair nowadays. This is our motivation to compare the latest
version of Dask with Spark.
% Methods (one sentence)
To evaluate the frameworks, we focus on their performance and their scheduler.

@@

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the
scheduler to perform more optimizations which makes it well suited for neuroimaging
analysis that often requires the usage of a pipeline with multiple steps. Though,
Spark's code base is in Scala which theoretically can lead to slow down in execution
due to a required serialization. Moreover, while Spark's API is flexible and allows
most implementations, it differs from the ones seen in Python's ecosystem.

% Dask
On the other hand, Dask was created with the purpose of natively parallelize Python
pipelines while keeping the syntax of familiar API from Python's scientific
ecosystem. However, Dask is still a young framework with work to be done; it's API
does not completely replicate the library it supports. While its lower-level API
allows the implementation of more complex algorithms it sacrifices a layer of
optimization.
% Related work
Previous work shows that Dask had significant overhead and was
hard to debug~\cite{Mehta:17}.
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework.
Therefore we think it is valuable to re-compare it with Spark.

%% Dask API
The Dask APIs we decide to use for our comparison are Dask bag, delayed and futures.
Dask bag offers an easy API to parallelize data. Dask delayed offer a lower-level API
that offers more flexibility; this is good to implement more complex tasks that do
not fit in the Dask bag framework. Dask futures is a real-time API. Both Dask bag and
delayed apply lazy evaluation to tasks while futures trigger them directly.

@@

\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality and lazy evaluation. Spark
abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
high-level API provides optimizations for transformations applied data. Since
Spark is written in Scala, this optimization is lessen because it requires a
serialization from Python to Java. The API offered by Spark is flexible and suits most
problems however it differs from Python ecosystem. The standalone scheduler is used for
our experiment. It has a LIFO policy.

\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
offered by Dask. Like Spark, it provides in-memory computing, data locality and lazy
evaluation however it does not have fault resilience. Fundamently, it is parallized
lists. Dask Bag parallelizes computations of large collection of generic python
objects~\cite{dask-bag}. Also, it offers optimized algorithms that mimic most of the
toolz~\cite{toolz} library. Our experiment use the Dask Distributed~\cite{dask-distributed}
scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
by finishing a branch of task computations before starting new ones.

\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
does not offer the optimized algorithm as the high-level APIs however it allows the
user to parallelize custom function that do not fit in the array, dataframe or
MapReduce framework. This is perfect for more complex function which could easily be
parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
fault resilience. The Dask Distributed scheduler is used.

\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
from Dask. Like the other Dask APIs, it benifits of in-memory computing and data
locatlity. However, unlike the other APIs, Dask Futures triggers computation
immediately. When the user tries to gather data if available it will return the
result otherwise it will block until the computation is done. Again, the Dask
Distributed scheduler is used.

\subsection{Dataset}
For the incrementation and histogram experiment the BigBrain~\cite{Amunts:13} is
used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
The data set store the intensity of each voxel and their position. This results in a
75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
perform our experiment we convert the chunks into the \textit{NifTI}~\cite{nifti}
format. Also, using \textit{sam}~\cite{sam} we combine the chunks and split them into
30, 125 and 750 chunks of equal size; each of 2.5GB, 0.6GB, and 0.1GB, respectively.

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and write the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all input are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimic the
behavior of analyzing multiple independent subjects in parallel.

@@

\subsubsection{\textbf{Histogram}}
Our second experiment consist to calculate the histogram of the intensity of the
voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
nfs, calculate inividual image frequency, aggregate the images frequency together,
and write the result to a file. In this experiment, the result depends on all the
input (see Fig.~\ref{fig:tg-histo}). This allows to study the behavior of the
frameworks when tasks are dependent. Also, this application requires data shuffling
thus inter-worker communication. We decided to omit Dask Futures as it does not had
any value over Dask Delayed in this application.

@@

\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
scaling of the frameworks in a cluster. Also, it helps in determining the level of
parallelism of the frameworks.

\textbf{Number of chunks:} Evaluating the frameworks on different amount of chunks
allows us to understand their behavior with different task size and IO requirements.

\textbf{Number of iterations:} We change the amount of iterations for the
incrementation~\ref{alg:incrementation} to study the effect of the job size.

\textbf{Sleep delay:} We modify the sleep delay in the
incrementation~\ref{alg:incrementation} to understand the behavior of task of
different length.

@@

\begin{table}[!b]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of chunks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

\subsection{Infrastructure}
All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
7.5.1804 as base operating system and have the Linux kernel version
3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
30GB 2666 MHz memory. The networks has a 10GB/s bandwidth. The filesystem is 2TB NFS v4.

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Experiment 1: Number of instances}
Figure~\ref{fig:inc_worker}(a) shows that the increase in performance is not
proportional to the number of workers which differ from what we would expect; it even
decrease when we reach 8 workers. This is explained by an increase in IO overhead
correlated to the number of workers (Figure~\ref{fig:inc_worker}(b)), as more worker
access the NFS simultaneously it increases the stress put on the system which results
in bottleneck, errors and slow down.

On the one hand, Dask seems better when using a single worker. On the other hand,
Spark makespan decreases faster for a higher level of parallelism. Dask is constantly
faster than Spark which makes us think it has a better scheduler.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of chunks}
In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
Spark has a 2GB limitation in the task size it can compute. Note from
Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
chunks. This is expected because the sleep delay is constant throughout this
experiment and there are more compute tasks when the number of chunks increases. We
also note that in Figure~\ref{fig:inc_tt_chunk} for 30 chunks the Dask bag API has a
much lower overhead time. This is because we only calculate the idle time of the used
core. Dask bag was only using one thread per block in comparison to Dask delayed and
Dask futures which was offloading the calculations on multiple threads of the same
worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
in more (smaller) chunks it saves IO time however it increases the overhead by a
similar amount; the inverse happens when we reduce the number of chunks.

@@

%% Iterations
\subsection{Experiment 1: Number of iterations}
In Figure~\ref{fig:inc_itr}, Spark and the Dask bag API are equivalent in terms of
performance. The Dask delayed API seems to perform well with a low amount of task but
grow faster the Spark and Dask bag. We think this is because Dask delayed often
schedule a block on a different thread of the same worker which causes small delays
every time due to inter-thread data communication. This is good when there are fewer
tasks as is make the IO between blocks go out of sync hence lowering the IO
bottleneck however as the number of tasks increases those small delays become
significant. The Dask futures API seems to outperform all the other APIs but this is
most likely because the tasks are not interdependent thus this application benefits
from the less optimal but faster scheduling brought by Dask futures.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
\end{figure}

%% Sleep time
\subsection{Experiment 1: Sleep delay}
As Figure~\ref{fig:inc_sleep}(b) and Table~\ref{tb:inc-sleep} shows Spark has the
smallest overhead all around and it is also the one that reduces the most with longer
task length. In term of overhead, Dask bag trailing slightly behind Spark although is
better than Dask delayed and Dask futures. The opposite happens for the IO time.
% Not sure what to conclude.

%%%% CHANGE NUMBER
\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-sleep}
    \centering
    \begin{tabular*}{\columnwidth}{llllll}
    \hline
                 & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular*}
 \end{table}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 8 instances}\label{fig:inc_sleep}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Table~\ref{tb:inc_base} and Figure~\ref{fig:inc_gantt}, the frameworks have
similar makespan however the time spend for each type of task differs significantly.
Spark tends to spend more time than other frameworks for IO however it has a much
lower overhead. Dask delayed and futures have the lowest IO time but a much higher
overhead. Dask bag stands in the middle for both IO and overhead. Compute time is
approximately the same for all frameworks.

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc_base}
    \centering
    \begin{tabular*}{\columnwidth}{llllll}
    \hline
                    & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular*}
 \end{table}

\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
\end{figure*}

@@

\section*{References}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,reference}

>>>>

% Approach
We propose to use Dask since it provides native Python parallelism to pipelines while
providing support to familiar API from the Python's scientific ecosystem.
Unfortunately, there are few comparisons between Spark and Dask; especially in
neuroimaging. Moreover, these studies were done when Dask was still an immature
framework which makes those studies unfair nowadays. This is our motivation to
compare the latest version of Dask with Spark.
% Methods (one sentence)
To evaluate the frameworks, we focus on their performance and their scheduler.

@@

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the
scheduler to perform more optimizations which makes it well suited for neuroimaging
analysis that often requires the usage of a pipeline with multiple steps. Though,
Spark's code base is in Scala which theoretically can lead to slow down in execution
due to a required serialization. Moreover, while Spark's API is flexible and allows
most implementations, it differs from the ones seen in the Python's ecosystem.

% Dask
On the other hand, Dask was created with the purpose of natively parallelize Python
pipelines while keeping the syntax of familiar API from the Python's scientific
ecosystem. However, Dask is still a young framework with work to be done; it's API
does not completely replicate the library it supports. While its lower-level API
allows the implementation of more complex algorithms it sacrifices a layer of
optimization.
% Related work
Previous work shows that Dask had significant overhead and was
hard to debug~\cite{Mehta:17}.
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework.
Therefore we think it is valuable to re-compare it with Spark.

%% Dask API
The Dask APIs we decide to use for our comparison are Dask Bag, Delayed and Futures.
Dask Bag offers an easy API to parallelize data. Dask Delayed offer a lower-level API
that offers more flexibility; this is good to implement more complex tasks that do
not fit in the Dask Bag framework. Dask Futures is a real-time API. Both Dask Bag and
Delayed apply lazy evaluation to tasks while Futures trigger them directly.

@@

\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality, and lazy evaluation. Spark
abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant.
Its high-level API provides optimizations for transformations applied data. Since
Spark is written in Scala, this optimization is lessened because it requires a
serialization from Python to Java. The API offered by Spark is flexible and suits
most problems however it differs from the Python ecosystem. The standalone scheduler
is used for our experiment. It has a LIFO policy.

\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
offered by Dask. Like Spark, it provides in-memory computing, data locality, and lazy
evaluation however it does not have fault resilience. Fundamentally, it is
parallelized lists. Dask Bag parallelizes computations across a large collection of
generic python objects~\cite{dask-bag}. Also, it offers optimized algorithms that
mimic most of the toolz~\cite{toolz} library. Our experiment uses the Dask
Distributed~\cite{dask-distributed} scheduler. It has a LIFO policy. In theory, it
should speed-up the pipeline execution by finishing a branch of task computations
before starting new ones.

\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
does not offer the optimized algorithm as the high-level APIs however it allows the
user to parallelize custom function that does not fit in the array, dataframe or
MapReduce framework. This is perfect for the more complex functions which could
easily be parallelized. Like Dask Bag, it offers in-memory computing, data locality
and lacks fault resilience. The Dask Distributed scheduler is used.

\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
from Dask. Like the other Dask APIs, it benefits of in-memory computing and data
locality. However, unlike the other APIs, Dask Futures triggers computation
immediately. When the user tries to gather data if available it will return the
result otherwise it will block until the computation is done. Again, the Dask
Distributed scheduler is used.

\subsection{Dataset}
For the incrementation and histogram experiment, the BigBrain~\cite{Amunts:13} is
used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
The dataset stores the intensity of each voxel and its position. This results in a
75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
perform our experiment we convert the chunks into the \textit{NifTI}~\cite{nifti}
format. Also, using \textit{sam}~\cite{sam} we combine the chunks and split them into
30, 125 and 750 chunks of equal size; each of 2.5GB, 0.6GB, and 0.1GB, respectively.

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and writes the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all inputs are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimics the
behavior of analyzing multiple independent subjects in parallel.

@@

\subsubsection{\textbf{Histogram}}
Our second experiment consists to calculate the histogram of the intensity of the
voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
NFS, calculate individual image frequency, aggregate the frequency of the images
together, and write the result to a file. In this experiment, the result depends on
all the input (see Fig.~\ref{fig:tg-histo}). This allows the study of the behavior of
the frameworks when tasks are dependent. Also, this application requires data
shuffling thus inter-worker communication. We decided to omit Dask Futures as it does
not have any value over Dask Delayed in this application.

@@

\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
scaling of the frameworks in a cluster. Also, it helps in determining the level of
the parallelism of the frameworks.

\textbf{Number of chunks:} Evaluating the frameworks on a different amount of chunks
allows us to understand their behavior with different task size and IO requirements.

\textbf{Number of iterations:} We change the number of iterations for the
incrementation~\ref{alg:incrementation} to study the effect of the job length.

\textbf{Sleep delay:} We modify the sleep delay in the
incrementation~\ref{alg:incrementation} to understand the behavior of different task
lengths.

@@

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of chunks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

\subsection{Infrastructure}
All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
7.5.1804 as the base operating system and have the Linux kernel version
3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
30GB 2666 MHz memory. The network has a 10GB/s bandwidth. The filesystem is 2TB NFS
v4.

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Experiment 1: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
the y-axis and the number of workers on the x-axis; the color represents the
different frameworks. Overall, there is no substantial makespan difference between
the frameworks. Also, the increase in performance is not proportional to the number
of workers; performance even decreases when we reach 8 workers.

In Figure~\ref{fig:inc_tt_worker}, the y-axis shows the total execution time of the
incrementation application. The hatch indicates that the time is spent for: read,
compute, write, and overhead. Again, the color represents the frameworks. The
computing time stays similar when the number of workers increases. However, the IO
time and overhead increase proportionally to the number of workers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of chunks}
In Figure~\ref{fig:inc_chunk}, Spark is not compared for 30 chunks. This is because
Spark has a 2GB limitation in the task size it can compute.

The Figure~\ref{fig:inc_ms_chunk} shows the makespan evolution when varying the
number of chunks. The compute time increase when there are more chunks. This is
because the sleep delay is constant throughout this experiment and there are more
compute tasks when the number of chunks increases.

In Figure~\ref{fig:inc_tt_chunk}, the total execution time of each function is shown.
For 30 chunks the Dask Bag API has a much lower overhead time. This is because we
only calculate the idle time of the used core. Dask Bag was only using one thread per
block in comparison to Dask Delayed and Dask Futures which was offloading a block
calculations on multiple threads of the same worker. Furthermore, the variance of the
overhead increases proportionally to the number of chunks. Also, the IO time reduces
with more chunks however the overhead time increase by a similar amount. This is not
observed for 30 chunks as the workers are not used at full capacity; i.e.\ some
threads are idle.

@@

%% Iterations
\subsection{Experiment 1: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
number of iterations. Spark and the Dask Bag API are equivalent in terms of
performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
it has a similar makespan to Spark and Dask Bag when the number of iterations
increases. Overall, Dask Futures seems to be faster but the difference is not
substantial.

In Figure~\ref{fig:inc_tt_itr}, the execution time of the function type is shown. The
compute time increase proportionally to the number of iterations. This is because
each computing task has a constant sleep delay. Also, the overhead increases when the
number of iterations (tasks) increases while the IO time decreases; however this is
not substantial.

% We think this is because Dask Delayed often
% schedule a block on a different thread of the same worker which causes small delays
% every time due to inter-thread data communication. This is good when there are fewer
% tasks as is make the IO between blocks go out of sync hence lowering the IO
% bottleneck however as the number of tasks increases those small delays become
% significant.

%  The Dask Futures API seems to outperform all the other APIs but this is
% most likely because the tasks are not interdependent thus this application benefits
% from the less optimal but faster scheduling brought by Dask Futures.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
\end{figure}

%% Sleep time
\subsection{Experiment 1: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
different sleep delays. Spark is initially slower than the Dask APIs, however, it is
because faster when increasing the sleep time. Also, within the Dask APIs, Dask Bag
is slower than the other two but it is not considerable.

Figure~\ref{fig:inc_tt_sleep} shows the execution time for each type of function. On
one hand, Spark has the smallest overhead. Table~\ref{tab:inc_sleep_overhead} shows
that it is substantial; especially with a 64 seconds sleep delay. On the other hand,
Spark has a high IO time. In contrast, Dask Delayed and Futures have the largest
overhead time but the smallest IO time. Then, Dask Bag stands in the middle for both
overhead and IO time.

\begin{table}[!t]
    \caption{Function time for the sleep experiment}
    \begin{subtable}[b]{\columnwidth}
        \renewcommand{\arraystretch}{1.3}
        \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
        \centering
        \begin{tabular}{lllll}
        \hline
                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
        Spark        & 14992  & 17076  & 16069   & 9769    \\
        Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
        Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
        Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
        \end{tabular}
    \end{subtable}
    \vskip 0.2cm
    \begin{subtable}[b]{\columnwidth}
        \renewcommand{\arraystretch}{1.3}
        \caption{IO time in second}\label{tab:inc_sleep_io}
        \centering
        \begin{tabular}{lllll}
        \hline
                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
        Spark        & 63622  & 57144  & 51070   & 57078   \\
        Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
        Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
        Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
        \end{tabular}
    \end{subtable}
 \end{table}

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 8 instances}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Table~\ref{tab:inc_base} and Figure, the frameworks have
similar makespan however the time spend for each type of task differs significantly.
Spark tends to spend more time than other frameworks for IO however it has a much
lower overhead. Dask Delayed and Futures have the lowest IO time but a much higher
overhead. Dask Bag stands in the middle for both IO and overhead. Compute time is
approximately the same for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
\end{figure*}

\begin{table}[!b]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in second: 125 chunks, 4 sec.\ sleep
    delay, 10 iterations, and 8 workers}\label{tab:inc_base}
    \centering
    \begin{tabular*}{\columnwidth}{llllll}
    \hline
                 & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular*}
 \end{table}

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,reference}
